---
title: "pre-class04"
author: "Gabriele Borg"
date: "October 3, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}
#create index
i = 1
#loop over index and print values
while(i<=50) {
  print(i)
  i = i+1
}
```



### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
#create sum_vec variable equal to zero
sum_vec<-0
#use the for loop this time to sum all integers from 0 to 1000
for(i in 0:1000) {
  sum_vec<-sum_vec + i
}
#print the sum vector
print(sum_vec)
```


B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
sum_even = 0
for (i in seq(0,1000, by=2)) {
  sum_even = sum_even + i
}
print(sum_even)
```


C. Now, repeat A and B WITHOUT using a loop.
```{r}
sum(seq(0,1000, by = 2))
```


### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
#create name vector
participant<-survey$participant
#loop over survey columns, replace wrong values and add it to list
for (i in 2:length(survey)) {
  temp <- survey[i]
  temp[(temp<0 | temp>5)] <- NA
  participant <- cbind(participant,temp)
}
#rename dataframe
survey.clean<- participant

#print dataframe
survey.clean
```


B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}

#loop through rows to compute no. of NAs
for (i in 1:length(survey.clean)){
  survey.clean$invalid.answers[i] <- sum(is.na(survey.clean[i,]))
}

#print dataframe
survey.clean

```

